Test a simpler version of veb-boost performance on ukbiobank data-2

### 2020-06-11 analysis/index.Rmd
plotly
Former-commit-id: a3bcd8a2db83a1e3728cc6c1216da4d0a67c0617

  * [Using UKbiobank data (chr22)](simulation-simpleveb-boost-ukbchr22.html)

@@

  * [Using UKbiobank data (chr22)](simulation-simpleveb-boost2-ukbchr22.html)

>>>>

  * [Using UKbiobank data (chr22)](https://users.rcc.uchicago.edu/~simingz/simulation-simpleveb-boost-ukbchr22.html)

@@

  * [Using UKbiobank data (chr22)](https://users.rcc.uchicago.edu/~simingz/simulation-simpleveb-boost2-ukbchr22.html)

### 2020-06-11 analysis/simulation-simpleveb-boost-ukbchr22.Rmd
plotly
Former-commit-id: a3bcd8a2db83a1e3728cc6c1216da4d0a67c0617

```{r plot functions}
library(mr.ash.alpha)
source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")

@@

plot_beta <- function(beta,beta.pm, ...){
  plot( beta, pch=19, col ="darkgreen", ...)
  points(beta.pm, pch =19, col = "red")
  legend("topright", legend=c("true beta", "posterior mean"),
       col=c("darkgreen", "red"), pch=19)
}

plot_pip <- function(indi, pip) {
  barplot(indi)
}

summary_mr.ash2 <- function(g.fit, s.fit, phenores){
  e.b <- rep(0, length(g.fit$beta))
  e.b[phenores$param$idx.cgene] <- phenores$param$e.beta
  cat("gene expression effect: \n")
  summary_mr.ash(g.fit)
  plot_beta(e.b, g.fit$beta)

  s.b <- rep(0, length(s.fit$beta))
  s.b[phenores$param$idx.cSNP] <- phenores$param$s.theta
  cat("snp effect: \n")
  summary_mr.ash(s.fit)
  plot_beta(s.b, s.fit$beta)
}

@@

MR.ASH results (start from gene):
```{r sim 1 gene}
load(paste0(outputdir, "mr.ash2_20200503-1-mr.ash2.expr-res.Rd"))
g.fit <- mr.ash2.fit$fit1
s.fit <-  mr.ash2.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 1 snp}
load(paste0(outputdir, "mr.ash2_20200503-1-mr.ash2.snp-res.Rd"))
g.fit <- mr.ash2.fit$fit2
s.fit <-  mr.ash2.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

@@

True parameters:
```{r para 2}
load(paste0(outputdir, "mr.ash2_20200503-2-pheno.Rd"))
readLines(paste0(outputdir, "param-20200503-2.R"))
```

MR.ASH results (start from gene):
```{r sim 2 gene}
load(paste0(outputdir, "mr.ash2_20200503-2-mr.ash2.expr-res.Rd"))
g.fit <- mr.ash2.fit$fit1
s.fit <-  mr.ash2.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 2 snp}
load(paste0(outputdir, "mr.ash2_20200503-2-mr.ash2.snp-res.Rd"))
g.fit <- mr.ash2.fit$fit2
s.fit <-  mr.ash2.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

@@

True parameters:
```{r para 3}
load(paste0(outputdir, "mr.ash2_20200503-3-pheno.Rd"))
readLines(paste0(outputdir, "param-20200503-3.R"))
```

MR.ASH results (start from gene):
```{r sim 3 gene}
load(paste0(outputdir, "mr.ash2_20200503-3-mr.ash2.expr-res.Rd"))
g.fit <- mr.ash2.fit$fit1
s.fit <-  mr.ash2.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 3 snp}
load(paste0(outputdir, "mr.ash2_20200503-3-mr.ash2.snp-res.Rd"))
g.fit <- mr.ash2.fit$fit2
s.fit <-  mr.ash2.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

@@

True parameters:
```{r para 4}
load(paste0(outputdir, "mr.ash2_20200503-4-pheno.Rd"))
readLines(paste0(outputdir, "param-20200503-4.R"))
```

MR.ASH results (start from gene):
```{r sim 4 gene}
load(paste0(outputdir, "mr.ash2_20200503-4-mr.ash2.expr-res.Rd"))
g.fit <- mr.ash2.fit$fit1
s.fit <-  mr.ash2.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 4 snp}
load(paste0(outputdir, "mr.ash2_20200503-4-mr.ash2.snp-res.Rd"))
g.fit <- mr.ash2.fit$fit2
s.fit <-  mr.ash2.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

>>>>

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})

source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")

@@

summary_mr.ash2 <- function(g.fit, s.fit, phenores, pipcut = 0.1){
  e.b <- rep(0, length(g.fit$beta))
  e.b[phenores$param$idx.cgene] <- phenores$param$e.beta
  cat("gene expression effect: \n")
  summary_mr.ash(g.fit)

  s.b <- rep(0, length(s.fit$beta))
  s.b[phenores$param$idx.cSNP] <- phenores$param$s.theta
  cat("snp effect: \n")
  summary_mr.ash(s.fit)
}

plot_pip <- function(dt){
  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

  fig <- fig %>% layout(yaxis = list(title = "PIP"))
  return(fig)
}

plot_beta <- function(dt){
  fig <- plot_ly(dt, x = ~p0, y = ~ TrueBeta, type = 'scatter', mode = 'markers', name ="Truth", text = ~paste("Name: ", name))
  fig <- fig %>% add_trace(y = ~ EstBeta, name = 'Estimated',mode = 'markers')

  fig <- fig %>% layout(yaxis = list(title = "beta"))
  return(fig)
}

plot_p <- function(dt){
  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  fig <- plot_ly(dt, x = ~p0, y = ~ PVALUE , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

  fig <- fig %>% layout(yaxis = list(title = "p value"))
  return(fig)
}

plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp){
  a <- read.table(mrashres.gene , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  p <- fread(gwasres.gene, header =T)
  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
  dt.gene <- merge(a, p, by = "name", all.x = T)

  a <- read.table(mrashres.snp , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  p <- fread(gwasres.snp, header =T)
  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
  dt.snp <- merge(a, p, by = "name", all.x = T)

  fig1 <- plot_pip(dt.gene)
  fig2 <- plot_beta(dt.gene)
  fig3 <- plot_p(dt.gene)
  fig4 <- plot_pip(dt.snp)
  fig5 <- plot_beta(dt.snp)
  fig6 <- plot_p(dt.snp)

  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, nrows=6, shareX = T, titleY = T)

  fig <- fig %>% layout(autosize = F, width = 600, height = 800, margin = 0.02,
    xaxis = list(title = "Genomic position"))

  fig <- fig %>% toWebGL()
  fig
}

@@

MR.ASH results (start from gene):
```{r sim 1 gene, out.width = "98%", out.height="95%"}
load(paste0(outputdir, "mr.ash2_20200503-1-mr.ash2.expr-res.Rd"))
g.fit <- mr.ash2.fit$fit1
s.fit <-  mr.ash2.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)

tag <- "mr.ash2_20200503-1"
plot_all(paste0(outputdir, tag, "-mr.ash2.expr-res.expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2.expr-res.snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
```

MR.ASH results (start from SNP):
Very similar to start from gene, therefore, not shown.
```{r sim 1 snp, eval=F}
load(paste0(outputdir, "mr.ash2_20200503-1-mr.ash2.snp-res.Rd"))
g.fit <- mr.ash2.fit$fit2
s.fit <-  mr.ash2.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)

tag <- "mr.ash2_20200503-1"
plot_all(paste0(outputdir, tag, "-mr.ash2.snp-res.expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2.snp-res.snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
```

@@

True parameters:
```{r para 2, eval=T}
load(paste0(outputdir, "mr.ash2_20200503-2-pheno.Rd"))
readLines(paste0(outputdir, "param-20200503-2.R"))
```

MR.ASH results (start from gene):
```{r sim 2 gene, eval=T, out.width = "98%", out.height="95%"}
load(paste0(outputdir, "mr.ash2_20200503-2-mr.ash2.expr-res.Rd"))
g.fit <- mr.ash2.fit$fit1
s.fit <-  mr.ash2.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)

tag <- "mr.ash2_20200503-2"
plot_all(paste0(outputdir, tag, "-mr.ash2.expr-res.expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2.expr-res.snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
```

@@

True parameters:
```{r para 3, eval=T}
load(paste0(outputdir, "mr.ash2_20200503-3-pheno.Rd"))
readLines(paste0(outputdir, "param-20200503-3.R"))
```

MR.ASH results (start from gene):
```{r sim 3 gene, eval=T, out.width = "98%", out.height="95%"}
load(paste0(outputdir, "mr.ash2_20200503-3-mr.ash2.expr-res.Rd"))
g.fit <- mr.ash2.fit$fit1
s.fit <-  mr.ash2.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)

tag <- "mr.ash2_20200503-3"
plot_all(paste0(outputdir, tag, "-mr.ash2.expr-res.expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2.expr-res.snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
```

@@

True parameters:
```{r para 4, eval=T}
load(paste0(outputdir, "mr.ash2_20200503-4-pheno.Rd"))
readLines(paste0(outputdir, "param-20200503-4.R"))
```

MR.ASH results (start from gene):
```{r sim 4 gene, eval=T, out.width = "98%", out.height="95%"}
load(paste0(outputdir, "mr.ash2_20200503-4-mr.ash2.expr-res.Rd"))
g.fit <- mr.ash2.fit$fit1
s.fit <-  mr.ash2.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)

tag <- "mr.ash2_20200503-4"
plot_all(paste0(outputdir, tag, "-mr.ash2.expr-res.expr.txt"), paste0(outputdir, tag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2.expr-res.snp.txt"), paste0(outputdir, tag, ".snpgwas.txt.gz"))
```

### 2020-06-11 analysis/simulation-simpleveb-boost2-ukbchr22.Rmd
plotly
Former-commit-id: a3bcd8a2db83a1e3728cc6c1216da4d0a67c0617

plot_beta <- function(beta,beta.pm, ...){
  plot( beta, pch=19, col ="darkgreen", ...)
  points(beta.pm, pch =19, col = "red")
  legend("topright", legend=c("true beta", "posterior mean"),
       col=c("darkgreen", "red"), pch=19)
}

plot_pip <- function(indi, pip, pipcut = 0.1) {
  x <- barplot(indi, main = "PIP")
  points(x, pip, pch=19, col="red")
  pred <- which(pip > pipcut)
  real <- which(indi==1)
  tp <- length(intersect(pred, real))/length(pred)
  fp <- 1 - tp
  fn <- length(setdiff(real, pred))/length(real)
  cat("false positive rate: ", fp, "\n")
  cat("false negative rate: ", fn, "\n")
}

summary_mr.ash2 <- function(g.fit, s.fit, phenores, pipcut = 0.1){
  e.b <- rep(0, length(g.fit$beta))
  e.b[phenores$param$idx.cgene] <- phenores$param$e.beta
  cat("gene expression effect: \n")
  summary_mr.ash(g.fit)
  plot_beta(e.b, g.fit$beta)
  indi <- rep(0, length(g.fit$beta))
  indi[phenores$param$idx.cgene] <- 1
  pip <- get_pip(g.fit)
  plot_pip(indi, pip, pipcut = pipcut)

  s.b <- rep(0, length(s.fit$beta))
  s.b[phenores$param$idx.cSNP] <- phenores$param$s.theta
  cat("snp effect: \n")
  summary_mr.ash(s.fit)
  plot_beta(s.b, s.fit$beta)
}

@@

MR.ASH2s results (start from gene):
```{r sim 1 gene}
load(paste0(outputdir, "20200527-1-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 1 snp}
load(paste0(outputdir, "20200527-1-mr.ash2s.snp-res.Rd"))
g.fit <- mr.ash2s.fit$fit2
s.fit <-  mr.ash2s.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

@@

MR.ASH results (start from gene):
```{r sim 2 gene}
load(paste0(outputdir, "20200527-2-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 2 snp}
load(paste0(outputdir, "20200527-2-mr.ash2s.snp-res.Rd"))
g.fit <- mr.ash2s.fit$fit2
s.fit <-  mr.ash2s.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

@@

MR.ASH results (start from gene):
```{r sim 3 gene}
load(paste0(outputdir, "20200527-3-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 3 snp}
load(paste0(outputdir, "20200527-3-mr.ash2s.snp-res.Rd"))
g.fit <- mr.ash2s.fit$fit2
s.fit <-  mr.ash2s.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

@@

MR.ASH results (start from gene):
```{r sim 4 gene}
load(paste0(outputdir, "20200527-4-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)
```

MR.ASH results (start from SNP):
```{r sim 4 snp}
load(paste0(outputdir, "20200527-4-mr.ash2s.snp-res.Rd"))
g.fit <- mr.ash2s.fit$fit2
s.fit <-  mr.ash2s.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)
```

>>>>

summary_mr.ash2 <- function(g.fit, s.fit, phenores, pipcut = 0.1){
  e.b <- rep(0, length(g.fit$beta))
  e.b[phenores$param$idx.cgene] <- phenores$param$e.beta
  cat("gene expression effect: \n")
  summary_mr.ash(g.fit)

  s.b <- rep(0, length(s.fit$beta))
  s.b[phenores$param$idx.cSNP] <- phenores$param$s.theta
  cat("snp effect: \n")
  summary_mr.ash(s.fit)
}

plot_pip <- function(dt){
  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

  fig <- fig %>% layout(yaxis = list(title = "PIP"))
  return(fig)
}

plot_beta <- function(dt){
  fig <- plot_ly(dt, x = ~p0, y = ~ TrueBeta, type = 'scatter', mode = 'markers', name ="Truth", text = ~paste("Name: ", name))
  fig <- fig %>% add_trace(y = ~ EstBeta, name = 'Estimated',mode = 'markers')

  fig <- fig %>% layout(yaxis = list(title = "beta"))
  return(fig)
}

plot_p <- function(dt){
  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  fig <- plot_ly(dt, x = ~p0, y = ~ PVALUE , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

  fig <- fig %>% layout(yaxis = list(title = "p value"))
  return(fig)
}

plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp){
  a <- read.table(mrashres.gene , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  p <- fread(gwasres.gene, header =T)
  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
  dt.gene <- merge(a, p, by = "name", all.x = T)

  a <- read.table(mrashres.snp , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  p <- fread(gwasres.snp, header =T)
  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
  dt.snp <- merge(a, p, by = "name", all.x = T)

  fig1 <- plot_pip(dt.gene)
  fig2 <- plot_beta(dt.gene)
  fig3 <- plot_p(dt.gene)
  fig4 <- plot_pip(dt.snp)
  fig5 <- plot_beta(dt.snp)
  fig6 <- plot_p(dt.snp)

  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, nrows=6, shareX = T, titleY = T)

  fig <- fig %>% layout(autosize = F, width = 600, height = 800, margin = 0.02,
    xaxis = list(title = "Genomic position"))

  fig <- fig %>% toWebGL()
  fig
}

@@

MR.ASH2s results (start from gene):
```{r sim 1 gene, out.width = "98%", out.height="95%"}
load(paste0(outputdir, "20200527-1-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)

tag <- "20200527-1"
simtag <- "mr.ash2_20200503-1"
plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
```

MR.ASH results (start from SNP):
Very similar to start from gene, therefore, not shown.
```{r sim 1 snp, eval = F}
load(paste0(outputdir, "20200527-1-mr.ash2s.snp-res.Rd"))
g.fit <- mr.ash2s.fit$fit2
s.fit <-  mr.ash2s.fit$fit1
summary_mr.ash2(g.fit, s.fit, phenores)

tag <- "20200527-1"
simtag <- "mr.ash2_20200503-1"
plot_all(paste0(outputdir, tag, "-mr.ash2s.snp-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2s.snp-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
```

@@

MR.ASH results (start from gene):
```{r sim 2 gene, out.width = "98%", out.height="95%"}
load(paste0(outputdir, "20200527-2-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)

tag <- "20200527-2"
simtag <- "mr.ash2_20200503-2"
plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
```

@@

MR.ASH results (start from gene):
```{r sim 3 gene, out.width = "98%", out.height="95%"}
load(paste0(outputdir, "20200527-3-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)

tag <- "20200527-3"
simtag <- "mr.ash2_20200503-3"
plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
```

@@

MR.ASH results (start from gene):
```{r sim 4 gene, out.width = "98%", out.height="95%"}
load(paste0(outputdir, "20200527-4-mr.ash2s.expr-res.Rd"))
g.fit <- mr.ash2s.fit$fit1
s.fit <-  mr.ash2s.fit$fit2
summary_mr.ash2(g.fit, s.fit, phenores)

tag <- "20200527-4"
simtag <- "mr.ash2_20200503-4"
plot_all(paste0(outputdir, tag, "-mr.ash2s.expr-res.expr.txt"), paste0(simdatadir, simtag, ".exprgwas.txt.gz"), 
         paste0(outputdir, tag, "-mr.ash2s.expr-res.snp.txt"), paste0(simdatadir, simtag, ".snpgwas.txt.gz"))
```
